Kernel Phase and Coronagraphy with Automatic Differentiation

### 2020-07-29 ms.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{Kernel Phase and Coronagraphy with Automatic Differentiation}
\shortauthors{B. J. S. Pope}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% self cal with closure phase
In the recovery of object phase information, among the most longstanding techniques is that of self-calibration with `closure phases' \citep[introduced in the context of radio astronomy by][]{jennison58}, in which phases are summed around three interferometric baselines which form a closing triangle. The phase error terms local to each subaperture cancel, so that three low-signal-to-noise (SNR) baseline phases deliver one high-SNR observable. In a non-redundant array (one in which no baseline vector is repeated between different pairs of subapertures) of sufficient size, a large number of closure phases can be obtained which anchor image reconstruction with great precision. `Closure amplitudes' which are resistant to fluctuations in input amplitude or gain can also be obtained by a similar construction using four telescopes \citep{twiss60,blackburn20}. The closure phase is the argument of a quantity called the triple product or `bispectrum', and it has recently been shown from the perspective of invariant theory in algebraic geometry that for a wide class of problems limited by phase noise, knowledge of the mean, power spectrum and bispectrum are necessary and sufficient for an optimal signal reconstruction \citep{bandeira17}.

@@

\citet{martinache10} introduced the idea of `kernel phase', suggesting that while for a redundant aperture closure phases no longer cancel out the contributions of aberrations $\phi$, nevertheless in general a left kernel operator $\mathbf{K}$ can be found via singular value decomposition (SVD) of $\mathbf{A}_\phi$ such that

@@

In order to investigate this discrepancy, we generated a grid of forwards models. Using the same code used to produce the kernel phase autodiff model, we generated input wavefronts with 20 Zernike modes of aberrations, each with random amplitudes normally-distributed with 15\,nm standard deviation, for a total peak-to-peak amplitude of 206\,nm. These were then shifted and added to simulate a range of binaries at different configurations and contrasts, and both the \citet{martinache20} pipeline and the autodiff pipeline were used to retrieve their parameters. The results are shown in Figure~\ref{method_comparison_ensemble}. The pipelines achieve broadly comparable performance on these simulations, with the \citet{martinache20} method being slightly better at closer separations and higher contrasts and autodiff performing marginally better at larger separations.

In order to investigate whether this was due to wavefront errors leaking in, we generated a cube of 1000 PSFs with 20 Zernike modes of aberrations, this time with 20\,nm standard deviation, for a total peak-to-peak wavefront amplitude of $295\pm 100$\,nm. We then retrieved the system parameters with \texttt{emcee} for the cube, taking as our observables the median kernel phases and the standard error of the mean over the ensemble (posterior in Figure~\ref{comparison_posterior_sim}; and also frame-by-frame using least squares (displayed in Figure~\ref{comparison_posterior_sim_cube}. For a simulated binary with fiducial $\rho = 125$\,mas, $\theta=86^\circ$, and $c = 25$, there is no sign of the large systematic difference between the methods as seen in real data. Both methods, whether frame by frame or on average, retrieve the true parameters at least comparably accurately. In \citet{martinache20} a large error term was added in quadrature in order to account for additional noise in the data unexplained by the diversity over exposures and uncalibrated by the kernel phase model. Whatever systematic this is may be treated differently by these two pipelines, and may need to be a subject for future work.

\begin{figure}
\plotone{method_comparison_ensemble.pdf}
\caption{Comparison of \citet[][blue stars]{martinache20} and autodiff (orange dots) kernel phase pipelines on simulated data, as a function of input contrast (left) and separation (right). When inspected, the samples with worse separation error after both pipelines are those at higher contrasts. Uncertainties are omitted for clarity, and because they are poorly defined in this context. \label{method_comparison_ensemble}}
\end{figure}

\begin{figure}
\plotone{comparison_posterior_sim.pdf}
\caption{Comparison of \citet[][blue linetype]{martinache20} and autodiff (black) kernel phase pipelines on simulated data with true parameters $\rho = 125$~mas, $\theta = 86^\circ$, $c=25$. We generate a cube of 1000 mock datasets using a 64-pix pupil model and Zernike aberrations as described in Section~\ref{sec:palomar}, and the observables used are the median and standard error of the mean of each kernel phase across this ensemble. Both datasets retrieve the true parameters very precisely, though with small systematic offsets, which are slightly larger for the autodiff method than for classic kernel phases. \label{comparison_posterior_sim}}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{Kernel Phase and Coronagraphy with Automatic Differentiation}
\shortauthors{B. J. S. Pope et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% self cal with closure phase
In the recovery of object phase information, among the most longstanding techniques is that of self-calibration with `closure phases' \citep[introduced in the context of radio astronomy by][]{jennison58}, in which phases are summed around three interferometric baselines which form a closing triangle. The phase error terms local to each subaperture cancel, so that three low-signal-to-noise (SNR) baseline phases deliver one high-SNR observable. In a non-redundant array (one in which no baseline vector is repeated between different pairs of subapertures) of sufficient size, a large number of closure phases can be obtained which anchor image reconstruction with great precision. `Closure amplitudes' which are resistant to fluctuations in input amplitude or gain can also be obtained by a similar construction using four telescopes \citep{twiss60,blackburn20}. The closure phase is the argument of a quantity called the triple product or `bispectrum', and it has recently been shown from the perspective of invariant theory in algebraic geometry that for a wide class of problems limited by phase noise, knowledge of the mean of a signal, its power spectrum, and its bispectrum are necessary and sufficient for an optimal signal reconstruction \citep{bandeira17}.

@@

\citet{martinache10} introduced the idea of `kernel phase' to generalize this idea to a redundant aperture. In this case, the propagation of phase noise from the pupil to baselines is no longer linear, but in the limit where aberrations are small it can be linearized. While in this case closure phases no longer exactly cancel out the contributions of aberrations $\phi$, nevertheless in general a left kernel operator $\mathbf{K}$ can be found via singular value decomposition (SVD) that annihilates $\mathbf{A}_\phi$, such that

@@

% In order to investigate this discrepancy, we generated a grid of forwards models. Using the same code used to produce the kernel phase autodiff model, we generated input wavefronts with 20 Zernike modes of aberrations, each with random amplitudes normally-distributed with 15\,nm standard deviation, for a total peak-to-peak amplitude of 206\,nm. These were then shifted and added to simulate a range of binaries at different configurations and contrasts, and both the \citet{martinache20} pipeline and the autodiff pipeline were used to retrieve their parameters. The results are shown in Figure~\ref{method_comparison_ensemble}. The pipelines achieve broadly comparable performance on these simulations, with the \citet{martinache20} method being slightly better at closer separations and higher contrasts and autodiff performing marginally better at larger separations.

In order to investigate the effect on both pipelines of wavefront errors leaking in, we generated a cube of 1000 PSFs with 20 Zernike modes of aberrations, this time with 20\,nm standard deviation, for a total peak-to-peak wavefront amplitude of $295\pm 100$\,nm. We then retrieved the system parameters with \texttt{emcee} for the cube, taking as our observables the median kernel phases and the standard error of the mean over the ensemble (posterior in Figure~\ref{comparison_posterior_sim}; and also frame-by-frame using least squares (displayed in Figure~\ref{comparison_posterior_sim_cube}. For a simulated binary with fiducial $\rho = 125$\,mas, $\theta=86^\circ$, and $c = 25$, there is no sign of the large systematic difference between the methods as seen in real data. Both methods, whether frame by frame or on average, retrieve the true parameters at least comparably accurately. In \citet{martinache20} a large error term was added in quadrature in order to account for additional noise in the data unexplained by the diversity over exposures and uncalibrated by the kernel phase model. Whatever systematic this is may be treated differently by these two pipelines, and may need to be a subject for future work.

% \begin{figure}
% \plotone{method_comparison_ensemble.pdf}
% \caption{Comparison of \citet[][blue stars]{martinache20} and autodiff (orange dots) kernel phase pipelines on simulated data, as a function of input contrast (left) and separation (right). When inspected, the samples with worse separation error after both pipelines are those at higher contrasts. Uncertainties are omitted for clarity, and because they are poorly defined in this context. \label{method_comparison_ensemble}}
% \end{figure}

% \begin{figure}
% \plotone{comparison_posterior_sim.pdf}
% \caption{Comparison of \citet[][blue linetype]{martinache20} and autodiff (black) kernel phase pipelines on simulated data with true parameters $\rho = 125$~mas, $\theta = 86^\circ$, $c=25$. We generate a cube of 1000 mock datasets using a 64-pix pupil model and Zernike aberrations as described in Section~\ref{sec:palomar}, and the observables used are the median and standard error of the mean of each kernel phase across this ensemble. Both datasets retrieve the true parameters very precisely, though with small systematic offsets, which are slightly larger for the autodiff method than for classic kernel phases. \label{comparison_posterior_sim}}
% \end{figure}
